P.berghei pir analysis

### 2021-09-13 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
source stacked_barchart function

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

@@

```{r stacked_barchart_fun}
stacked_barchart <- function(melted_tpm,
                             threshold = 1,
                             stack_group_name = 'Geneid',
                             sd_bars = FALSE){
  #Have to use sym and !! in order to get mutate to work 
  # (converting the stack_group_name to a symbol and then unquoting in the mutate function)
  stack_group_name <- rlang::sym(paste(stack_group_name))

  tpm_data <- mutate(melted_tpm, 
                     stack_group = (!!stack_group_name)) %>% 
    mutate(tpm = ifelse(tpm > threshold, tpm, 0)) %>% #Filtering step for low value pir genes.
    group_by(stage, SorL, stack_group, sample) %>% 
    summarise(sum_tpm = sum(tpm)) %>% 
    group_by(SorL, stack_group, stage) %>%
    summarise(mean_sum_tpm = mean(sum_tpm),
              sd_sum_tpm = sd(sum_tpm)) %>% 
    arrange(desc(SorL),
            mean_sum_tpm) %>% 
    ungroup() %>% 
    mutate(order = row_number())

  tpm_data_SorL_stackgenes <- 
    #Order by SorL and then by mean_sum_tpm, then convert stack_group to a factor so ggplot orders correctly
    arrange(tpm_data,
            stage,
            desc(SorL), 
            mean_sum_tpm) %>% 
    mutate(stage = factor(stage,
                          levels = ordering(unique(tpm_data$stage),
                                            order_vector = r_order_vector)),
           stage_num = as.integer(
             factor(stage,
                    levels = ordering(unique(tpm_data$stage),
                                      order_vector = r_order_vector))
           )) %>% 
    group_by(stage) %>% 
    mutate(stacked_mean = cumsum(mean_sum_tpm),
           sd_upper = cumsum(mean_sum_tpm) + cumsum(sd_sum_tpm),
           sd_lower = ifelse(cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm) < 0,
                             0, 
                             cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm)))

  tpm_data_SorL_stackgenes$stack_group_order <- seq_len(length.out = nrow(tpm_data_SorL_stackgenes))

  stacked_max <- tpm_data %>% 
    ungroup %>% 
    group_by(stage) %>% 
    dplyr::summarise(max_tpm = sum(mean_sum_tpm)) %>% 
    .$max_tpm %>% 
    max

  max_tpm <- c(
    seq(
      0,
      mround(
        stacked_max,
        500),
      length.out = 11)
  )

  stage_vec <- tpm_data_SorL_stackgenes$stage
  col_group <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'zzAsex' ,
                         str_detect(stage_vec, pattern = 'Liv') ~ 'zzLiv',
                         str_detect(stage_vec, pattern = 'Gam') ~ 'zzGam',
                         str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'zzMos')
  tpm_data_SorL_stackgenes$stage_col <- col_group
  fills <- c('#F6B9AD','#96d3ff','#f0d4e3','#a7e9af') 
  #Note that this order is influenced by sort(col_group)

  # Note that this code does not work with ggplot 3.3.5, but it does work with 3.3.3. 
  #  Some issue with the colours for the geom_rect background.

  y <- ggplot(data = tpm_data_SorL_stackgenes, 
              aes(
                x = stage_num,
                y = mean_sum_tpm, 
                fill = SorL
              )) +
    geom_rect(
      aes(
        xmin = stage_num - .7,
        xmax = stage_num + .7,
        ymin = -Inf,
        ymax = Inf,
        fill = stage_col
      ),
      alpha = .5
    ) +
    geom_bar(position = position_stack(),
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
    {if(sd_bars)list(
      geom_errorbar(aes(ymax = sd_upper,
                        ymin = sd_lower),
                    position = position_dodge(width = .85),
                    size = .25),
        geom_point(aes(y = stacked_mean), 
                   position = position_dodge(width = .85),
                   show.legend = FALSE,
                   size = .25)
    )}+ 
    scale_fill_manual(
      breaks = unique(tpm_data_SorL_stackgenes$SorL),
      values = c(
        SorL_col_num,
        fills
      )
    ) +
    scale_color_manual('black') + 
    scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) +
    scale_x_continuous(
      breaks = tpm_data_SorL_stackgenes$stage_num %>% unique(),
      labels = tpm_data_SorL_stackgenes$stage %>% unique(),
      expand = c(0, 0)
    ) +
    theme(
      axis.text.x = element_text(angle = 90, 
                                 vjust = .5,
                                 size = 10),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 10),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
    ) +
    xlab('stage') + 
    facet_grid(.~stage, 
               scales = "free_x",
               space = "free_x") +
    scale_x_discrete(expand=c(0,0)) +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 10),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=8, 
                                  face = 'bold', 
                                  angle = 90),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      ),
      panel.spacing.x = unit(0, "null")
    )

  g <- ggplot_gtable(ggplot_build(y))
  strip_t <- which(grepl('strip-t', g$layout$name))
  stage_vec_strip <- sort(unique(tpm_data_SorL_stackgenes$stage))
  fills <- case_when(str_detect(stage_vec_strip, pattern = 'Asex') ~ 'coral1' ,
                     str_detect(stage_vec_strip, pattern = 'Liv') ~ 'plum2',
                     str_detect(stage_vec_strip, pattern = 'Gam') ~ 'lightskyblue1',
                     str_detect(stage_vec_strip, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
  k <- 1
  for (i in strip_t) {
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }

  grid.newpage()
  grid.draw(g)
  return(g)
}
```

>>>>

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
source('../rfunctions/stacked_barchart_fun.R')
```

@@
